Alien species

### 2023-03-22 DRAFT_alien_species.Rmd
masking example

4. Then, look at [Median Summer Temperature] for how to aggregate spatially tp accounting areas (regions)

This workflow should be synchronized with the [slitasje](#slitasje-og-kjørespor) indicator.

>>>>

4. Then, look at [Median Summer Temperature] for how to aggregate spatially to accounting areas (regions)

This workflow should be synchronized with the [slitasje](#slitasje-og-kjørespor) indicator. I will attempt a little proof of concept below.

## Masking with an ecosystem delineation - example

Let's try steps 2 and 3 on a reduced data set. I'll work with wetlands in Gran municipality.

```{r}
gran <- wetland_alien_extr %>%
  filter(municipalityNumber == "3446")
```

```{r, fig.cap="Alien species indicator over wetlands in Gran municipality."}
tm_shape(gran)+
  tm_polygons(col = "w_mean",
    title="Alien species indicator - wetlands",
    palette = myCol,
     style="fixed",
    breaks = seq(0,1,.1))+
  tm_layout(legend.outside = T,
            title = "Gran municipality")
```

Then I can import and crop the ecosystem delineation map.

```{r, eval=F}
file <- "P:/41201785_okologisk_tilstand_2022_2023/data/Myrmodell/myrmodell90pros.tif"
EDM <- stars::read_stars(file, proxy=F)
```

The EDM is in UTM32 when we actually want to have it in UTM33. But transforming it takes too long, so I will transform the indicator map instead.

```{r, eval=FALSE}
gran <- st_transform(gran, st_crs(EDM))
```

```{r, eval=F}
mire_gran <- st_crop(EDM, gran)
saveRDS(mire_gran, paste0(pData, "cache/mire_gran.rds"))
```

```{r, include=FALSE}
mire_gran <- readRDS(paste0(pData, "cache/mire_gran.rds"))
```

```{r, fig.cap="Wetlands in Gran municipality"}
ggplot()+
  geom_stars(data = st_downsample(mire_gran, 10))
```

Then I rasterize the indicator map to match the EDM. This takes a while, even for just one municipality. I'm not sure this would work for the entire country.

```{r, eval=F}
gran_rast <- st_rasterize(gran, template = mire_gran)  
saveRDS(gran_rast, paste0(pData, "cache/gran_rast.rds"))
```

```{r, include=FALSE}
gran_rast <- readRDS(paste0(pData, "cache/gran_rast.rds"))
```

```{r}
gran_outline <- muni %>%
  filter(kommunenummer == "3446")
```

```{r, fig.cap="Indictor values (alien species, extrapolated and area weighted values) in Gran municipality"}
tm_shape(st_downsample(gran_rast, 10))+
  tm_raster(col="w_mean")+
  tm_shape(gran_outline)+
  tm_borders(col="red")
```

Then i mask

```{r}
gran_rast_masked <- gran_rast
gran_rast_masked[mire_gran == 0] <- NA
```

```{r, fig.cap="Indictor values (alien species, extrapolated and area weighted values) in Gran municipality"}
tm_shape(st_downsample(gran_rast_masked, 10))+
  tm_raster(col="w_mean",
            palette = "red")+
  tm_shape(gran_outline)+
  tm_borders(col="red")
```

From here it is easy to get zonal statistics using `exactextratr`. If I have this kind of map for all of Norway I can get the mean value for the regions (defined via a polygon set) and it will of course be area weighted.
